### 2024-04-19 build/44b7303602dca5cee397811da222885d.tex
initial commit; 0

\documentclass[crop,dvisvgm]{standalone}
\usepackage {amsmath,tikz-cd}
\begin{document}
\begin {tikzcd} 
    1  \ar [dd, no head] & 4  \ar [d, no head] & & 9  \ar [d, no head] &
   \\ & 5  \ar [d, no head] & & 10  \ar [dl, no head]  \ar [dr, no head] &
   \\ 2  \ar [dd, no head]  \ar [dr, no head] & 6  \ar [d, no head] & 11  \ar [ddr, no head] & & 12  \ar [ddl, no head]
   \\ & 7  \ar [d, no head]  \ar [drr, no head] & & &
   \\ 3  \ar [dddddrr, no head] & 8 & & 13  \ar [d, no head] &
   \\  & & & 14  \ar [d, no head] &
   \\  & & & 15  \ar [d, no head] &
   \\  & & & 16  \ar [d, no head]  \ar [dr, no head]  \ar [ddl, no head] &
   \\  & & & 17  \ar [dl, no head, dashed] & 18
   \\  & & 19 & &
   \end {tikzcd}
\end{document}