### 2024-04-20 MSRI-16-1989-79.Rmd
starting 2

::: {.rmenv #chunk-2.2 title="2.2" latex="{2.2}"}
For an abelian variety $A$ over $\QQ$, we denote by $T(A)\otimes\QQ$ the motive $\HH_1^\mot(A)$ that is dual to $\HH_\mot^1(A)$, and by $T(A)$ the motive with integer coefficients defined by the integer structure $\HH_1(A(\CC),\ZZ)\subset\HH_1(A(\CC),\QQ)=(T(A)\otimes\QQ)_\BB$.
\oldpage{106}The functor $A\mapsto T(A)$ is fully faithful: from $T(A)$ we can recover $\Lie(A)=T(A)_\DR/F^0$ and the complex torus
$$
  A(\CC)
  = T(A)_\BB\setminus\Lie(A)
  = T(A)_\BB\setminus T(A)_\DR\otimes\CC/F^0.
$$
The complex torus $A(\CC)$ determines the abelian variety $A_\CC$ over $\CC$ induced by $A$, and the $\QQ$-structure is uniquely determined by that of the Lie algebra.

An abelian scheme $A$ over $S$ similarly defines a smooth motive with integer coefficients $T(A)$ over $S$.
:::

::: {.rmenv #chunk-2.3 title="2.3" latex="{2.3}"}
Recall that a \em{smooth $1$-motive} $X$ over a scheme $S$ consists of

a. a group scheme $L$ over $S$ that, locally, for the étale topology, is a constant group scheme defined by a free $\ZZ$-module of finite type;
  an abelian scheme $A$ over $S$, and a torus $T$ over $S$;
b. an extension $E$ of $A$ by $T$, and a morphism $\bar{u}\colon L\to A$;
c. a morphism $u\colon L\to E$ lifting $\bar{u}$.

We write $X=[L\xrightarrow{u}E]$.

A $1$-motive $X$ over $\QQ$ defines a motive over $\QQ$ with integer coefficients $T(X)$ (cf. [[D3](D3), Section 10, but the crystalline aspect is missing]), and the functor $X\mapsto T(X)$ is fully faithful (cf. [[D3](D3), both 10.1.3 and 2.2]).

For $X=[\ZZ\to0]$, $T(X)$ is the unit motive $\ZZ(0)$.
For $X=[0\to\Gm]$, $T(X)$ is the Tate motive $\ZZ(1)$.
For an abelian variety $A$ and $X=[0\to A]$, $T(X)=T(A)$.
Of course, here, as in \ref{minus-three-points-2.2}, we can take more general bases than $\Spec(\QQ)$.
:::

::: {.rmenv #chunk-2.4 title="2.4" latex="{2.4}"}
I conjecture that the set of motives with integer coefficients of the form $T(X)$ for some $1$-motive $X$ is stable under extensions.
If $T'$ is a motive with integer coefficients, with $T'\otimes\QQ\simto T(X)\otimes\QQ$, then $T'$ is of the form $T(X')$ with $X'$ isogenous to $X$.
The conjecture is thus equivalent to the claim that the set of motives $T(X)\otimes\QQ$, for $1$-motives $X$, is stable under extensions.
The word "conjecture" is an abuse of terminology, since the statement itself is not precise.
What is conjectured is that every realisation system \ref{minus-three-points-1.9} (or \ref{minus-three-points-1.24}, over $S$) that is an extension $T(X)$ by $T(Y)$ (for $1$-motives $X$ and $Y$), and "natural", "coming from geometry", is isomorphic to that defined by a $1$-motive $Z$ that is an extension of $X$ by $Y$.

\oldpage{107}
A point $a$ of an abelian variety $A$ over $\QQ$ defines a $1$-motive $[\ZZ\xrightarrow{u}A]$ with $u(1)=a$.
The motive $T([\ZZ\xrightarrow{u}A])$ is an extension of $\ZZ(0)$ by $T(A)$, and the conjecture, applied to $\ZZ(0)$ and $T(A)$, implies that
$$
  A(\QQ) \simto \Ext^1(\ZZ(0),T(A))
$$
or, equivalently,
$$
  A(\QQ)\otimes\QQ \simto \Ext^1(\QQ(0),T(A)\otimes\QQ)
$$
(where $\Ext^1$ is in the abelian category of motives).

More generally, if $E$ is an extension of an abelian variety by a torus, we want
$$
  E(\QQ) \simto \Ext^1(\ZZ(0),T(E))
$$
and similarly for more general bases that $\Spec(\QQ)$.

The case $E=\Gm$ is particularly interesting: for every smooth scheme $S$ over $\Spec(\ZZ)$, we want, in the category of motives with integer coefficients over $S$,

$$
  \Gamma(S,\OO_S^\times) \simto \Ext^1(\ZZ(0),\ZZ(1)).
\startverb\tag{2.4.1}\stopverb
$$

:::

:::: {.rmenv #chunk-2.5 title="2.5" latex="{2.5}"}
Let $M$ be a realisation system \ref{minus-three-points-1.9}.
We spell out what a realisation system $E$ given by an extension of $\QQ(0)$ by $M$ is, when $M$ is of weight $\leqslant-1$ (i.e. $M=W_{-1}(M)$).

::: {.trmenv #M1P title="(M1)P" latex="{(M1)P}"}
$E_\BB$ is an extension $M_\BB\to E_\BB\xrightarrow{u}\QQ$ of $\QQ$ by $M_\BB$.
Its data is equivalent to that of $P_\BB\coloneqq u^{-1}(1)$, endowed with its structure as a torsor (principal homogeneous space) for $M_\BB$, i.e. $m\in M_\BB$ acts on $P_\BB$ via $p\mapsto p+m$.
:::

::: {.trmenv #M234P title="(M2,3,4)P" latex="{(M2,3,4)P}"}
Similarly, the data of $E_\DR$ (resp. $E_\AAf$, $E_\crisp$) is equivalent to that of a torsor $P_\DR$ (resp. $P_\AAf$, $P_\crisp$) for $M_\DR$ (resp. $M_\AAf$, $M_\crisp$).
:::

::: {.trmenv #M5P title="(M5)P" latex="{(M5)P}"}
The comparison morphisms for $E$ can be identified with comparison morphisms $P_\BB\otimes\CC\simto P_\DR\otimes\CC$, $P_\BB\otimes\AAf\simto P_\AAf$, and $P_\DR\otimes\QQ_p\simto P_\crisp$ that are compatible with the analogous morphisms for $M$.

\subtree{
  \taxon{notation}
  \p{
    #{\otimes}, see \ref{minus-three-points-0.4}.
  }
}
:::

::: {.trmenv #M6P title="(M6)P" latex="{(M6)P}"}
Since $M$ is, by hypothesis, of weight $\leqslant-1$, and $\QQ(0)$ is of pure weight $0$, the weight filtration of $E$ is determined by that of $M$, i.e. $W_0(E)=E$, and $W_i(E)=W_i(E)$ for $i<0$.
:::

::: {.trmenv #M7P title="(M7)P" latex="{(M7)P}"}
\oldpage{108}
The data of $F_\infty$ for $E$ is equivalent to that of an involution of $P_\BB$ that is compatible with that of $M_\BB$.
Axiom [(AM2)](minus-three-points-am2) for $E$ is that this involution is induced by the $\CC$-antilinear involution of $P_{\BB,\CC}=\PP_{\DR,\CC}$ with fixed points $P_{\DR,\RR}$.
:::

::: {.trmenv #M8P title="(M8)P" latex="{(M8)P}"}
The data of the Hodge filtration of $E_\DR$ is equivalent to that of the intersection $F^0P_\DR$ of $F^0E_\DR$ over $P_\DR$, which is a torsor for $F^0M_\DR$: for $p>0$, $F^pE_\DR=F^pM_\DR$; for $p=0$, $F^0E_\DR$ is generated by $F^0P_\DR$; and for $p<0$, $F^pE_\DR$ is the sum of $F^0E_\DR$ and $F^pM_\DR$.
In other words, the data [(M8)](minus-three-points-m8) for $E_\DR$ is that of a reduction of the structure group of $P_\DR$, from $M_\DR$ to $F^0M_\DR$.

The data of [(M1)](minus-three-points-m1), [(M2)](minus-three-points-m2), $\comp_{\DR/\BB}$, and [(M8)](minus-three-points-m8) are equivalent to the data of a torsor $F^0P_\DR$ for $F^0M_\DR$ and, in the $M_\DR\otimes\CC=M_\BB\otimes\CC$ torsor $P_{\DR,\CC}$ which is induced by $F^0M_\DR\to M_\DR\otimes\CC$, a $\QQ$-structure $P_\BB\subset P_{\DR,\CC}$ that is an $M_\BB$ torsor.
:::

::: {.trmenv #M9P title="(M9)P" latex="{(M9)P}"}
The data of the action of $\Gal(\bar{\QQ}/\QQ)$ on $E_\AAf$ is equivalent to that of an action of $\Gal(\bar{\QQ}/\QQ)$ on $P_\AAf$, which makes $P_\AAf$ an equivariant $M_\AAf$-torsor.
:::

::: {.trmenv #M10P title="(M10)P" latex="{(M10)P}"}
Similarly, the automorphism $\phi_p$ of $E_\DR\otimes\QQ_p$ corresponds to an automorphism $\phi_p$ of $P_\DR\otimes\QQ_p$ such that
$$
  \phi_p(x+m) = \phi_p(x) + \phi_p(m).
$$
:::

We leave to the reader the task of translating axioms [(AM1)](minus-three-points-am1) to [(AM5)](minus-three-points-am5) for $P$.
We note only that [(AM1)](minus-three-points-am1) and [(AM4)](minus-three-points-am4) follow from the same axioms for $M$, and that [(AM2)](minus-three-points-am2) and [(AM5)](minus-three-points-am5) can be expressed at best as functoriality in $\CC$, cf. \ref{minus-three-points-1.14}.
::::

::: {.rmenv #chunk-2.6 title="2.6 Definition" latex="{2.6 Definition}"}
—

i. A system of torsors, [(M1)P](minus-three-points-m1P) to [(M10)P](minus-three-points-m10P) in \ref{minus-three-points-2.5}, that satisfies the conditions required in order to define an extension $E$ of $\QQ(0)$ by a realisation system $M$ is called a \em{torsor in realisation systems}.

ii. If $M$ is a motive, then the torsor $P$ for $M$ is \em{motivic} (or a \em{torsor in motives}) if the extension $E$ is a motive.
:::

:::: {.rmenv #chunk-2.7 title="2.7" latex="{2.7}"}
Let $M$ be a realisation system with integer coefficients.
An \em{extension} $E$ of $\ZZ(0)$ by $M$ is an extension $E\otimes\QQ$ of $\QQ(0)$ by $M\otimes\QQ$, endowed with an integer structure such that the sequence $0\to M_\BB\to E_\BB\to\ZZ\to0$ is exact.

As in \ref{minus-three-points-2.5}, such an extension is described by a system of torsors.
\oldpage{109}As in \ref{minus-three-points-1.24} and \ref{minus-three-points-1.25}, this corresponds to modifying \ref{minus-three-points-2.5} as follows:

::: {.trmenv #M1PZ title="(M1)PZ" latex="{(M1)P$\ZZ$}"}
$P_\BB$ is an $M_\BB$-torsor.
:::

::: {.trmenv #M3t9PZ title="(M3,9)PZ" latex="{(M3,9)P$\ZZ$}"}
$P_\ell$ is, for each $\ell$, a $\Gal(\bar{\QQ}/\QQ)$-equivariant $M_\ell$-torsor.
It is equivalent to give a projective system of equivariant $M_{\ZZ/(n)}$-torsors $P_{\ZZ/(n)}$, or their projective limit $P_{\hZZ}$, which is an equivariant $M_{\hZZ}$-torsor.
:::

::: {.trmenv #M5PZ title="(M5)PZ" latex="{(M5)P$\ZZ$}"}
Replace $\comp_{\AAf,\BB}$ by $\comp_{\ell,\BB}$ for each $\ell$, or $\comp_{\ZZ/(n),\BB}$, or $\comp_{\hZZ,\BB}$.
:::

Such a system of torsors, which defines an extension of $\ZZ(0)$ by $M$, is also called a \em{torsor for $M$}.
It is said to be \em{motivic} if the corresponding extension is, i.e. if the torsor for $M\otimes\QQ$ that it defines is motivic.
::::

:::: {.rmenv #chunk-2.8 title="2.8" latex="{2.8}"}
We leave to the reader the task of understanding the constructions with bases $S$ more general than $\Spec(\QQ)$.
We note only that, in the de Rham realisation, $P_\DR$ is an $M_\DR$-torsor on $S_\QQ$.
It is endowed with a restriction of the structure group $F^0P_\DR\subset P_\DR$ to $F^0M_\DR$.
It is also endowed with an integrable connection $\nabla$.
Along each section $p$ of $P_\DR$, the vertical tangent space can be identified with $M_\DR$;
the connection $\nabla$ is thus a morphism
$$
  \nabla\colon P_\DR\to\Omega^1\otimes M_\DR.
$$
It satisfies $\nabla(p+m)=\nabla p+\nabla m$.
Transversality becomes the following:

::: {.trmenv #chunk-2.8.1 title="2.8.1" latex="{2.8.1}"}
For a local section $p$ of $F^0P_\DR$, $\nabla p\in\Omega^1\otimes F^{-1}(M_\DR)$.
:::
::::

::: {.rmenv #chunk-2.9 title="2.9" latex="{2.9}"}
We explicitly describe the torsor in motives $K(x)$ for $\ZZ(1)$ (the \em{Kummer torsor}) corresponding to $x\in\QQ^\times$ as in \ref{minus-three-points-2.4.1}.

The $\ZZ(1)_\BB$-torsor $K(x)_\BB$ is the torsor for $2\pi i\ZZ$ of logarithms of $x$.
Since $F^0\ZZ(1)_\DR=0$, we have $K(x)_\DR=\ZZ(1)_\DR=\QQ$.
The $\ZZ/(n)(1)$-torsor $K(x)_{\ZZ/(n)}$ is the torsor of $n$-th roots of unity of $x$.
The transition morphisms $K(x)/_{\ZZ(nm)}\to K(x)_{\ZZ/(n)}$ are given by $y\mapsto y^m$.
The Galois action is evident.

The morphism $\comp_{\DR,\BB}$ is induced by the inclusion of $K(x)_\BB$ into $\CC$.
The morphism $\comp_{\ZZ/(n),\BB}$ is induced by $K(x)_\BB\to K(x)_{\ZZ/(n)}$ given by $z\mapsto\exp(z/n)$.

The crystalline Frobenius $\phi_p\colon K(x)_\DR\otimes\QQ_p\to K(x)_\DR\otimes\QQ_p$, i.e. $\QQ_p\to\QQ_p$, is defined for $x$ a unit at $p$.
We denote by $\log$ the $p$-adic logarithm, which is characterised by
$$
  \phi_p^{-1}(0) = \log(x^{1-p}).
$$
:::

::: {.rmenv #chunk-2.10 title="2.10" latex="{2.10}"}
\oldpage{110}On $\Gm=\Spec\ZZ[q,q^{-1}]$, the universal section $q$ of $\Gm$ also defines a torsor in motives $K(q)$ for $\ZZ(1)$, on the base $\Gm$.
We now describe it.

The $\ZZ(1)_\BB$-torsor $K(q)_\BB$ on $C^\times$ is the local system of logarithms of $q$.
Since $F^0\ZZ(1)_\DR=0$, $K(q)_\DR$ is trivial: $K(n)_\DR=\OO$ (on ${\Gm}_\QQ$).
It is endowed with the integrable connection $\nabla\colon K(q)_\DR\to\Omega^1\otimes\ZZ(1)_\DR=\Omega^1$ given by
$$
  \nabla u = \dd u - \frac{\dd q}{q}.
$$
The $\ZZ/(n)(1)$-torsor $K(q)_{\ZZ/(n)}$ is that of the $n$-th roots of $q$.
This is a locally constant sheaf on $(\Gm\otimes\ZZ[1/n])_\et$.
Transition and comparison morphisms are defined as in \ref{minus-three-points-2.9}.

We now explain what the $F$-crystal structure is at $p$.
Let $\hGm$ be the completion of $\Gm$ along its fibre of characteristic $p$, and $\hGm\otimes\QQ_p$ the corresponding rigid analytic space.
For a Frobenius lifting $\widetilde{F}\colon\hGm\to\hGm$, the crystalline Frobenius is a morphism $\phi_p[\widetilde{F}]$, over $\hGm\otimes\QQ_p$, from $\widetilde{F}^\bullet K(q)_\DR$ to $K(q)_\DR$ that is horizontal and compatible with the torsor structures.
If we change $\widetilde{F}$ to $\widetilde{F}'$, then $\widetilde{F}$ and $\widetilde{F}'$ have the same reduction modulo $p$, the connection $\nabla$ gives an isomorphism of $\widetilde{F}^\bullet K(q)_\DR$ with $\widetilde{F}'^{\bullet} K(q)_\DR$, and the diagram
$$
  \begin{CD}
    \widetilde{F}^\bullet K(q)_\DR @>{\phi_p[\widetilde{F}]}>> K(q)_\DR
  \\@| @|
  \\\widetilde{F}'^{\bullet} K(q)_\DR @>>{\phi_p[\widetilde{F}']}> K(q)_\DR
  \end{CD}
$$
commutes.

In terms of $\phi_p^{-1}\colon K(q)_\DR\to\widetilde{F}'^{\bullet} K(q)_\DR$, and via the identifications $K(q)_\DR=\ZZ(1)_\DR=\OO$, whence $\widetilde{F}'^{\bullet}=\OO$, the compatibility with the torsor structures and with the connections can be written as
$$
  \begin{aligned}
    \phi_p^{-1}(k+m)
    &= \phi_p^{-1}(k) + pm
  \\\dd\phi_p^{-1}(0)
    &= \widetilde{F}^\bullet\left(\frac{\dd q}{q}\right) - p\frac{\dd q}{q}.
  \end{aligned}
$$
When we change the Frobenius lifting, we have
$$
  \phi_p^{-1}[\widetilde{F}'](0) - \phi_p^{-1}[\widetilde{F}](0)
  = \log(\widetilde{F}'/\widetilde{F}).
$$
For $\widetilde{F}\colon q\mapsto q^p$, we have
$$
  \phi_p^{-1}[q\mapsto q^p](0) = 0.
$$
\oldpage{111}For $x\in\QQ$ a unit at $p$, and $\widetilde{F}$ such that $\widetilde{F}(x)=x$, we thus have $\phi_p^{-1}[\widetilde{F}](0)=\log(x/x^p)$ at $x$, as required by \ref{minus-three-points-2.9}
:::

# Torsors for $\ZZ(n)$ {#section-3}

\oldpage{112}The study of the fundamental group of $\PP^1\setminus\{0,1,\infty\}$ reveals, for each $k\geqslant2$, a torsor $P_{1,k}$ for $\ZZ(k)$ on $\Spec(\ZZ)$.
We will describe these torsors in this section.
The relation with $\pi_1$ will be explained in \fref{minus-three-points-16} (see \fref{minus-three-points-16.23}) and the calculations to justify this will be given in \fref{minus-three-points-19}.
We will then describe the torsors $P_{d,k}$ for $\ZZ(k)$ on $\Spec(\QQ(\sqrt[d]{1}))$, given by an analogous construction \fref{minus-three-points-16.28}.

The sum of the $\P_{d,k}$ and of $(-1)^k$ times its symmetry under $\zeta\mapsto\zeta^{-1}$ is torsion.
This is essentially equivalent to the adelic integral formulas for the value of the Dirichlet $L$ functions at negative integers: \fref{minus-three-points-3.6}, \fref{minus-three-points-3.8}, \fref{minus-three-points-3.9}, and \fref{minus-three-points-3.14}.
A geometric interpretation, which allows us to reprove these formulas, will be given in \fref{minus-three-points-18}.

A geometric interpretation of the relations of distributions \fref{minus-three-points-3.13} will be given in \fref{minus-three-points-17}.

::: {.rmenv #chunk-3.1 title="3.1" latex="{3.1}"}
We denote by $P$ the torsor $P_{1,k}$ that we are going to define.
At the same time, we will define a trivialisation $u_\BB\in P_\BB$ of its Betti realisation.
Since $F^0\ZZ(k)_\DR=0$, we necessarily have that $P_\DR=\ZZ(k)_\DR=\QQ$.
The Betti structure is
$$
  \begin{aligned}
    P_\BB
    &= u_\BB + (2\pi i)^k\ZZ \subset \PP_{\DR,\CC} = \CC
    \quad\text{with}
  \\u_\BB
    &= -(k-1)!\zeta(k)
  \end{aligned}
\startverb\tag{3.1.1k}\stopverb
$$
where $\zeta$ is the Riemann $\zeta$ function.

Let $N$ be an integer; we will describe $P_{\ZZ/(N)}$.
For $z\in\CC$ with $z\neq1$, let $T_N(z)\coloneqq K(1-z)_{\ZZ/(N)}$ be the $\ZZ/N(1)$-torsor of $N$-th roots of $1-z$.
For $z=1$, we set $T_N(1)=K(1/N)_{\ZZ/(N)}$.
This convention ensures that the sum of the $T_N(\alpha)$ for $\alpha^N=1$ is trivialised: we have
$$
  \frac{1}{N}\prod(1-\alpha)
  = 1.
$$

Let $\alpha$ be an $N$-th root of $1$.
It defines $\alpha\colon\ZZ/N\to\ZZ/N(1)$ by $a\mapsto a\cdot\alpha$.
We thus obtain $\alpha^{\otimes(k-1)}\colon\ZZ/N(1)\to\ZZ/N(k)$.
Note that, for $\alpha=1$, this is the zero map.
We denote by $\alpha^{\otimes(k-1)}T(\alpha)$ the $\ZZ/N(k)$-torsor induced from $T(\alpha)$ by applying $\alpha^{\otimes(k-1)}$.
We have
$$
  P_{\ZZ/(N)}
  \coloneqq \sum_{\alpha^N=1} \alpha^{\otimes(k-1)}T_N(\alpha).
\startverb\tag{3.1.2k}\stopverb
$$
\oldpage{113}It is equivalent to sum over the $\alpha\neq1$.
This definition still makes sense in any algebraic closure $\overline{k}$ of $\QQ$, and the action of $\Gal(\bar{\QQ}/\QQ)$ is given by transport of structure.

We now define the transition morphisms $P_{\ZZ/(NM)}\to P_{\ZZ/(N)}$.
We have
$$
  \prod_{u^M=z} (1-u)
  = 1-z
\startverb\tag{3.1.3}\stopverb
$$
whence, for $z\neq0,1$, we obtain an isomorphism
$$
  T_N(z)
  = \prod_{u^M=z} T_N(u).
\startverb\tag{3.1.4}\stopverb
$$
If $\beta^N=1$ and $\alpha^M=\beta$, then the reduction $\mod N$ of $\alpha\colon\ZZ/NM\to\ZZ/NM(1)$ is $\beta\colon\ZZ/N\to\ZZ/N(1)$, and that of $T_{NM}(\alpha)$ is $T_N(\alpha)$.
From this, we obtain an isomorphism
$$
  \begin{aligned}
    P_{\ZZ/(NM)}\mod N
    &= \sum_{\beta^N=1} \sum_{\alpha^M=\beta} \alpha^{\otimes(k-1)}T_{NM}(\alpha) \mod N
  \\&= \sum_{\beta^N=1} \beta^{\otimes(k-1)} \sum_{\alpha^M=\beta} T_N(\alpha)
  \\&= \sum_{\beta^N=1} \beta^{\otimes(k-1)} T_N(\beta)
  \\&= P_{\ZZ/(N)}.
  \end{aligned}
$$
This defines the desired transition morphism.

For non-real negative $z$, we define the *principal determination* of $z\mapsto z^{1/N}$, and denote by $z^{1/N}$, that which takes the value of $1$ at $1$.
The morphism of torsors $\comp_{\ZZ/(N),\BB}$ sends $u_\BB$ to the point $u$ of $P_{\ZZ/(N)}$ defined by the $(1-\alpha)^{1/N}\in T(\alpha)$:
$$
  u_\BB
  \mapsto u
  = \sum \alpha^{\otimes(k-1)} ((1-\alpha)^{1/N}).
$$

If $z$ is not real and $\geqslant1$, then we have principal determinations
$$
  \prod_{u^M=z} (1-u)^{1/N}
  = (1-z)^{1/N}
\startverb\tag{3.1.5}\stopverb
$$
as we verify by analytic continuation from $z=0$, taking (3.1.3) into account.
It follows from (3.1.5) that the $\comp_{\ZZ/(N),\BB}$ are compatible with the transition morphisms.
:::

::: {.rmenv #chunk-3.2 title="3.2" latex="{3.2}"}
\oldpage{114}Some preliminaries, before defining $\phi_p$.
The $k$-logarithm is defined for $|z|\leqslant1$ by
$$
  \ell_k(z)
  = \sum z^n/n^k.
\startverb\tag{3.2.1}\stopverb
$$
The $\zeta(k)$ in the formula giving $P_\BB$ appears in our approach as the value at $1$ of $\ell_k$.
Let $p$ be a prime number and consider the formal series
$$
  \ell_k^{(p)}
  = \sum_{p\nmid n} z^n/n^k.
\startverb\tag{3.2.2}\stopverb
$$
This is the Taylor series expansion of a rigid analytic function on the projective line over $\QQ_p$ with the open disc of radius $1$ centred at $1$ removed.
Indeed, we have
$$
  \ell_k^{(p)}(z)
  = \lim_N \frac{1}{1-z^{p^N}} \sum_{n=1, p\nmid n}^{p^N-1} z^n/n^k.
\startverb\tag{3.2.3}\stopverb
$$
From (3.2.3) we get the identity
$$
  \ell_k^{(p)}(1/z)
  = -(-1)^k\ell_k^{(p)}(z).
\startverb\tag{3.2.4}\stopverb
$$

A formal series calculation gives, for $(N,p)=1$,
$$
  N^{k-1}\sum_{u^N=z} \ell_k^{(p)}(u)
  = \ell_k^{(p)}(z).
\startverb\tag{3.2.5}\stopverb
$$
This leads us to define the "principal value" at $1$
$$
  \vp\ell_k^{(p)}(1)
  \coloneqq \frac{N^{k-1}}{1-N^{k-1}} \sum_{\alpha^N=1,\alpha\neq1} \ell_k^{(p)}(\alpha).
\startverb\tag{3.2.6}\stopverb
$$
This number is independent of the choice of $N$ coprime to $p$.
By (3.2.4), it is zero for $k$ even.

With this notation, the crystalline Frobenius $\phi_p$ is given by
$$
  \phi_p^{-1}(0)
  = (k-1)!p^k\cdot\vp\ell_k^{(p)}(1).
\startverb\tag{3.2.7k}\stopverb
$$
:::

::: {.rmenv #chunk-3.3 title="Remark 3.3" latex="{Remark 3.3}"}
\oldpage{115}We have described the $\ZZ/(N)(k)$-torsor $P_{\ZZ/(N)}$ as a sum of torsors indexed by the $N$-th roots of $1$.
If $N$ is a power of a prime number $\ell$, say $N=\ell^n$, then we can again, thanks to (3.1.4), write it as a sum indexed by the set $\mu(N)^\times$ of primitive $N$-th roots of $1$.
Indeed, we have isomorphisms
$$
  \begin{aligned}
    P_{\ZZ/(N)}
    &= \sum_{\alpha^N=1}\ldots
  \\&= \sum_{a=1}^n \sum_{\mu(\ell^a)^\times} \ldots
  \\&= \sum_{a=1}^n \sum_{\alpha\in\mu(\ell^a)^\times}\alpha^{\otimes(k-1)} \sum_{\beta^{\ell^{n-a}}=\alpha} T_N(\beta).
  \end{aligned}
$$
In this sum, $\beta$ runs over, for each $a$, all the primitive $N$-th roots of $1$, whence
$$
  \begin{aligned}
    P_{\ZZ/(N)}
    &= \sum_{\beta\in\mu(N)^\times}\beta^{\otimes(k-1)}T_N(\beta) \cdot \sum_{a=1}^n\ell^{(n-a)(k-1)}
  \\&= \frac{1}{1-\ell^{k-1}} \sum_{\beta\in\mu(N)^\times}\beta^{\otimes(k-1)}T_N(\beta).
  \end{aligned}
$$
The cohomology class of $P_{\ZZ/(N)}$ (for $N=\ell^n$) in $\HH^1(\Spec(\QQ),\ZZ/(N)(k))$ thus admits the following description.
Let $\QQ(\zeta)$ be the cyclotomic field generated by a primitive $N$-th root of $1$.
In $\HH^1(\Spec(\QQ(\zeta)),\ZZ/N(1))$, let $\delta(1-\zeta)$ be the Kummer coboundary of $1-\zeta$.
On $\Spec(\QQ(\zeta))$, $\zeta$ identifies the $\ZZ/N(a)$ with one another, and we can think of $\delta(1-\zeta)$ as a class with values in $\ZZ/N(k)$.
More precisely, we consider $\zeta^{\otimes(k-1)}\cup\partial(1-\zeta)\in\HH^1(\Spec(\QQ(\zeta)),\ZZ/N(k))$.
We have
$$
  \cl(P_{\ZZ/\ell^n})
  = \frac{1}{1-\ell^{k-1}} \Tr_{\QQ(\zeta)/\QQ}(\zeta^{\otimes(k-1)}\cup\partial(1-\zeta))
$$
and the class of $P_\ell$ is the limit of these classes.
:::

::: {.rmenv #chunk-3.4 title="Remark 3.4" latex="{Remark 3.4}"}
We denote by $\zeta^{(p)}(s)$ the $\zeta$ function with its Eulerian factor at $p$ removed:
$$
  \zeta^{(p)}(s)
  \coloneqq \sum_{p\nmid n} 1/n^s.
$$
We know that the values of $\zeta^{(p)}(s)$ on the negative integers are rational and admit a $p$-adic interpolation.
More precisely,

a. Let $\CC_p$ be the completion of an algebraic closure of $\QQ_p$.
  Then the characters $\chi\colon\ZZ_p^\times\to\CC_p^\times$ are the points of a rigid analytic space.

b. \oldpage{116}There exists exactly one rigid analytic function $\zeta_p^{(p)}(\chi)$, with a simple pole at $\chi=(x\mapsto x^{-1})$, such that
  $$
    \zeta_p^{(p)}(x\mapsto x^{-1})
    = \zeta^{(p)}(n)\in\QQ
  $$
  for every integer $n<0$.

We set, for $n\in\ZZ$ with $n\neq1$,
$$
  \zeta_p^{(p)}(n)
  \coloneqq \zeta_p^{(p)}(x\mapsto x^{-1}).
$$
This notation differs from that of [@C] or [@Ko]: our $\zeta_p^{(p)}$ is not a continuous $p$-adic function of the *integer* $n$.
Instead, it is a continuous function of the *character* $x\mapsto x^n$, itself a continuous function of $n$ if and only if we restrict $n$ to an arithmetic progression of difference $p-1$.
We have
$$
  \zeta_p^{(p)}(n)
  = \lim_a \zeta^{(p)}(n-(p-1)p^a).
$$
The trivial zeros of $\zeta$ force $\zeta_p^{(p)}(n)=0$ for even $n$, i.e. for $x\mapsto x^{-n}$ an even character.
More generally, $\zeta_p^{(p)}(\chi)=0$ for even $\chi$.

With this notation, we have (cf. [@C; @Ko])
$$
  \vp \ell_k^{(p)}(1)
  = \zeta_p^{(p)}(k)
$$
and (3.2.7) can be rewritten as
$$
  \phi_p^{-1}(0)
  = (k-1)! p^k\zeta_p^{(p)}(k).
\startverb\tag{3.4.1k}\stopverb
$$
:::

::: {.itenv #chunk-3.5 title="Proposition 3.5" latex="{Proposition 3.5}"}
For $k$ even, the torsor $P_{1,k}$ is torsion, with order equal to the denominator of $\frac{1}{2}\zeta(1-k)$.
:::

::: {.rmenv #chunk-3.6 title="Paraphrasing 3.6" latex="{Paraphrasing 3.6}"}
Let $P$ be a torsor for a realisation system $M$ with integer coefficients.
The fact that $P$ is $n$-torsion means more than "every realisation is $n$-torsion".
Let $\frac{1}{n}M$ be the realisation system $M\otimes\QQ$, endowed with the integer structure $\frac{1}{n}M_\BB$.
It is isomorphic to $M$, by an isomorphism that sends $M\hookrightarrow\frac{1}{n}M$ by $n\colon M\to M$.
The fact that $P$ is $n$-torsion thus implies that it becomes trivial under $M\to\frac{1}{2}M$.
For $M$ of weight $\leqslant-1$, this implies:

a. \oldpage{117}$P_\BB+\frac{1}{n}M_\BB\cap F^0 P_\DR\otimes\CC\neq\phi$ (the intersection in $P_\BB\otimes\CC\simto P_\DR\otimes\CC$).
  The point of intersection $x$ is then unique, since $M_\BB\cap F^0(M_\DR)\otimes\CC=0$.
b. $x\in F^0 P_{DR}$, i.e. is defined over $\QQ$ for the $\QQ$-structure $\DR$.
c. The image of $x$ under $\comp_{\AAf,B}$ is fixed by Galois.
d. The image of $x$ under $\comp_{\crisp,\DR}$ is fixed by $\phi_p$.

Suppose that $F^0 M_\DR=0$.
Then $P_\DR=M_\DR$, and we can identify $P_\BB$ with a lateral class of $M_\BB\subset M_\DR\otimes\CC$.
Condition (a) then becomes $P_\BB\subset\frac{1}{n}M_\BB$.
We have $x=0$, and (b) is automatic.
We can write (d) as $\phi_p(0)=0$.
Let $p\in P_\BB$.
Denote by $p_M$ its image in $\frac{1}{n}M_\BB$, and set $p^\wedge=\comp_{\hZZ,\BB}(p)\in P_{\hZZ}$ and $(p_M)^\wedge=\comp_{\AAf,\BB}(p_M)\in M_\AAf$.
Condition (c) can then be written as
$$
  \sigma p^\wedge - p^\wedge
  = (\sigma-1)(p_M)^\wedge.
\startverb\tag{3.6.1}\stopverb
$$

Let $M=\ZZ(k)$ with $k\geqslant1$.
Let $\sigma\in\Gal(\bar{\QQ}/\QQ)$ with image $c$ in $\hZZ^\times$, i.e. $\sigma(\zeta)=\zeta^c$ for $\zeta$ a root of unity.
Condition (3.6.1) can then be written as
$$
  \sigma p^\wedge - p^\wedge
  = (c^k-1)(p_M)^\wedge.
\startverb\tag{3.6.2}\stopverb
$$
Modulo $N$, we obtain: let $c$ be an integer coprime to $Nn$, with $\sigma$ inducing $\zeta\mapsto\zeta^c$ on the $(Nn)$-th roots of unity.
Then $(c^k-1)p_M\in M_\BB$, and
$$
  \comp_{\hZZ,\BB}((c^k-1)p_M)
  \equiv \sigma p^\wedge - p^\wedge \mod{N}.
\startverb\tag{3.6.3}\stopverb
$$

In [(3.5)](#chunk-3.5), $k$ is even, and we can take $p=-(k-1)!\zeta(k)$, and (3.6.3) is a Galois formula for $-(c^k-1)(k-1)!\zeta(k)$ modulo $N(2\pi i)^k\ZZ$.
We recover a known formula expressing $\zeta$ on the negative integers as an adelic integral.
:::

::: {.rmenv title="Remark" latex="{Remark}"}
The above description shows that, if $P$ is of finite order, then its order can be read from its Hodge realisation.
:::

::: {.rmenv #chunk-3.7 title="3.7 (Proof of 3.5)" latex="{3.7 (Proof of 3.5)}"}
Let
:::